---
title: "project_final"
output: html_document
date: "2025-05-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(tidyr)
library(forcats)
library(reshape2)
library(tidyverse)
library(broom)
library(sjPlot)
library(sandwich)
library(lmtest)
library(lme4)
library(purrr)
library(stringr)
```

# Import Data


```{r}

# This table contains the result of the interview, the household 
# factor and other household characteristics.
household_data <- read.csv(unz("DATABASE.zip", "CS_HOGARES.csv"))

# This table contains the socio-demographic characteristics of the 
# household members, health status and utilization of health services, among others.
individual_data <- read.csv(unz("DATABASE.zip", "CS_RESIDENTES.csv"))

# The table includes information on variables related to the health 
# status of adults, such as obesity, hypertension, depressive symptoms, 
# diabetes mellitus, reproductive health, among others, which were 
# collected in the health questionnaire for adults (20 years and older). 
adult_diseases_data <- read.csv(unz("DATABASE.zip","CS_ADULTOS.csv"))

# The table includes variables from section IV of the household 
# questionnaire and those corresponding to the questionnaire of health 
# service users, these variables capture situations regarding the use of 
# ambulatory services, accessibility and quality of health services, 
# medicines, among others.
health_quality_data <- read.csv(unz("DATABASE.zip","CS_SERV_SALUD.csv"))
```

# Descriptives


```{r}
state_names <- c(
  "1" = "Aguascalientes", "2" = "Baja California", "3" = "Baja California Sur",
  "4" = "Campeche", "5" = "Coahuila", "6" = "Colima", "7" = "Chiapas",
  "8" = "Chihuahua", "9" = "CDMX", "10" = "Durango", "11" = "Guanajuato",
  "12" = "Guerrero", "13" = "Hidalgo", "14" = "Jalisco", "15" = "México",
  "16" = "Michoacán", "17" = "Morelos", "18" = "Nayarit", "19" = "Nuevo León",
  "20" = "Oaxaca", "21" = "Puebla", "22" = "Querétaro", "23" = "Quintana Roo",
  "24" = "San Luis Potosí", "25" = "Sinaloa", "26" = "Sonora", "27" = "Tabasco",
  "28" = "Tamaulipas", "29" = "Tlaxcala", "30" = "Veracruz", "31" = "Yucatán",
  "32" = "Zacatecas"
)

# Count households by state and exclude NA values
household_counts <- household_data %>%
  filter(!is.na(ENT)) %>%
  count(ENT, name = "num_households") %>%
  mutate(state_name = state_names[as.character(ENT)])

# Plot
ggplot(household_counts, aes(x = reorder(state_name, -num_households), y = num_households)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Number of Households by State",
       x = "State",
       y = "Number of Households") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
## UPDATING THE DATABASES WITHOUT TOP 5 STATES
```{r}
household_data <- household_data %>%
  filter(ENT != 2 & ENT != 9 & ENT != 14 & ENT != 19 & ENT != 22)

individual_data <- individual_data %>%
  filter(ENT != 2 & ENT != 9 & ENT != 14 & ENT != 19 & ENT != 22)

adult_diseases_data <- adult_diseases_data %>%
  filter(ENT != 2 & ENT != 9 & ENT != 14 & ENT != 19 & ENT != 22)

health_quality_data <- health_quality_data %>%
  filter(ENT != 2 & ENT != 9 & ENT != 14 & ENT != 19 & ENT != 22)

```

## DESCRIPTIVE STATISTICS 
From individual_data, "do you have health problems, where do you usually go for care?"
```{r}
labels <- c(
  "P3_9_01" = "Social Security (IMSS)",
  "P3_9_02" = "ISSSTE",
  "P3_9_03" = "State-level ISSSTE",
  "P3_9_04" = "Pemex",
  "P3_9_05" = "Defensa",
  "P3_9_06" = "Navy",
  "P3_9_07" = "Seguro Popular",
  "P3_9_08" = "IMSS PROSPERA",
  "P3_9_09" = "Private insurance",
  "P3_9_10" = "Other institution",
  "P3_9_11" = "Not affiliated",
  "P3_9_77" = "Other",
  "P3_9_12" = "None",
  "P3_9_99" = "Doesn't know"
)

# Select the columns of interest
binary_vars <- c("P3_9_01", "P3_9_02", "P3_9_03", "P3_9_04", "P3_9_05",
                 "P3_9_06", "P3_9_07", "P3_9_08", "P3_9_09", "P3_9_10",
                 "P3_9_11", "P3_9_77", "P3_9_12", "P3_9_99")

# Summarize counts of 1s
counts_df <- individual_data %>%
  select(all_of(binary_vars)) %>%
  summarise(across(everything(), ~ sum(. == 1, na.rm = TRUE))) %>%
  pivot_longer(cols = everything(), names_to = "code", values_to = "count") %>%
  mutate(label = labels[code]) %>%
  arrange(desc(count))

ggplot(counts_df, aes(x = reorder(label, count), y = count)) +
  geom_col(fill = "skyblue") +
  coord_flip() +  # horizontal for long labels
  labs(
    title = "Where do you usually go for care",
    x = "Service",
    y = "Number of People"
  ) +
  theme_minimal(base_size = 12)
```

People that is entitled to or has access to these medical services

```{r}
labels <- c(
  "1" = "Social Security (IMSS)",
  "2" = "ISSSTE",
  "3" = "State-level ISSSTE",
  "4" = "Pemex",
  "5" = "Defensa",
  "6" = "Navy",
  "7" = "Seguro Popular",
  "8" = "IMSS PROSPERA",
  "9" = "Private insurance",
  "10" = "Other institution",
  "11" = "Not affiliated",
  "99" = "Doesn't know"
)

# Create data frame with counts
access_df <- individual_data %>%
  count(P3_10_OPC1) %>%
  filter(P3_10_OPC1 %in% names(labels)) %>%
  mutate(service = labels[as.character(P3_10_OPC1)]) %>%
  arrange(desc(n))  # sort by frequency

# Plot
ggplot(access_df, aes(x = reorder(service, n), y = n)) +
  geom_col(fill = "skyblue") +
  coord_flip() +  # horizontal bars = better for long labels
  labs(
    title = "Access to Medical Services",
    x = "Healthcare Service",
    y = "Number of People"
  ) +
  theme_minimal(base_size = 12)

```

## PROPORTION OF PEOPLE BY STRATA
Here we can see that the highest strata also goes to the highest quality health providers
```{r}
# Create proportional data
proportional_data <- individual_data %>%
  filter(!is.na(P3_10_OPC1) & P3_10_OPC1 != 99) %>%
  group_by(ESTRATO, P3_10_OPC1) %>%
  summarise(count = n(), .groups = "drop") %>%
  mutate(service = labels[as.character(P3_10_OPC1)]) %>%
  filter(!is.na(service)) %>%
  group_by(ESTRATO) %>%
  mutate(percentage = count / sum(count) * 100)

# Create 100% stacked bar chart
ggplot(proportional_data, aes(x = factor(ESTRATO), y = percentage, fill = service)) +
  geom_bar(stat = "identity", position = "fill") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  labs(
    title = "Proportional Distribution of Health Services by Economic Strata (By Quartile)",
    x = "Economic Strata (ESTRATO)",
    y = "Percentage",
    fill = "Health Service"
  ) +
  theme_minimal() +
  theme(legend.position = "right")

```


## CREATING QUALITY INDEX BY PROVIDER
```{r}
# Reverse code negatively oriented Likert variables
health_quality_data <- health_quality_data %>%
  mutate(
    P4_6_rc = ifelse(P4_6 %in% 1:5, 6 - as.numeric(as.character(P4_6)), NA),
    P7_7_rc = ifelse(P7_7 %in% 1:5, 6 - as.numeric(as.character(P7_7)), NA)
  )

# Recode Yes/No variables: 1 = Yes = 1, 2 = No = 0, others = NA
yes_no_vars <- c("P7_2", "P8_8_1", "P8_8_2", "P8_8_3")
health_quality_data <- health_quality_data %>%
  mutate(across(all_of(yes_no_vars), ~ case_when(
    . == 1 ~ 1,
    . == 2 ~ 0,
    TRUE ~ NA_real_
  )))

# Convert Likert variables to numeric (1 to 5)
likert_vars <- c(
  "P3_5", "P8_2", "P8_3", "P8_4",
  "P8_5_1", "P8_5_2", "P8_5_3", "P8_5_4", "P8_5_5", "P8_5_6", "P8_6"
)
health_quality_data <- health_quality_data %>%
  mutate(across(all_of(likert_vars), ~ ifelse(. %in% 1:5, as.numeric(as.character(.)), NA_real_)))

# Define scoring rules
likert_5_vars <- c(
  "P3_5", "P4_6_rc", "P7_7_rc", "P8_2", "P8_3", "P8_4",
  "P8_5_1", "P8_5_2", "P8_5_3", "P8_5_4", "P8_5_5", "P8_5_6", "P8_6"
)
binary_vars <- c("P7_2", "P8_8_1", "P8_8_2", "P8_8_3")

# Calculate raw score, max possible score, and normalized index
health_quality_data <- health_quality_data %>%
  mutate(
    quality_score_raw = rowSums(across(all_of(c(likert_5_vars, binary_vars))), na.rm = TRUE),
    quality_score_max = rowSums(across(all_of(likert_5_vars), ~ ifelse(!is.na(.), 5, 0))) +
      rowSums(across(all_of(binary_vars), ~ ifelse(!is.na(.), 1, 0))),
    quality_score = ifelse(quality_score_max > 0, quality_score_raw / quality_score_max, NA_real_)
  )
```
 
Compare quality score across providers.
```{r}
# Compare quality_score across providers

# Recode provider labels
labels <- c(
  "1" = "Social Security (IMSS)",
  "2" = "ISSSTE",
  "3" = "State-level ISSSTE",
  "4" = "Pemex",
  "5" = "Defensa",
  "6" = "Navy",
  "7" = "Seguro Popular",
  "8" = "IMSS PROSPERA",
  "9" = "Private insurance",
  "10" = "Other institution",
  "11" = "Not affiliated",
  "99" = "Doesn't know"
)

health_quality_data <- health_quality_data %>%
  mutate(P3_7_label = recode(as.character(P3_7), !!!labels))

# Summarize normalized quality_score by provider
provider_summary <- health_quality_data %>%
  filter(!is.na(quality_score), !is.na(P3_7_label)) %>%
  group_by(P3_7_label) %>%
  summarise(
    avg_quality = mean(quality_score, na.rm = TRUE),
    n = n()
  ) %>%
  arrange(desc(avg_quality))

# Plot average quality score by provider
ggplot(provider_summary, aes(x = fct_reorder(P3_7_label, avg_quality), y = avg_quality)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    x = "Healthcare Provider",
    y = "Average Quality Score (0–1)",
    title = "Average Patient-Reported Quality by Healthcare Provider"
  ) +
  theme_minimal()

```

Heatmap.
```{r}
# Labels for heatmap dimensions
dimension_labels <- c(
  P3_5     = "Quality of service",
  P4_6_rc  = "Waiting time",
  P7_7_rc  = "Facility conditions",
  P8_2     = "Ease of getting an appointment",
  P8_3     = "Access outside regular hours",
  P8_4     = "Phone access to doctor",
  P8_5_1   = "Knows your history",
  P8_5_2   = "Lets you ask questions",
  P8_5_3   = "Spends enough time",
  P8_5_4   = "Involves you in decisions",
  P8_5_5   = "Explains clearly",
  P8_5_6   = "Coordinates specialist care",
  P8_6     = "Overall care rating"
)

# Transform to long format and label dimensions
long_data <- health_quality_data %>%
  select(P3_7_label, all_of(names(dimension_labels))) %>%
  pivot_longer(-P3_7_label, names_to = "dimension", values_to = "value") %>%
  mutate(dimension = recode(dimension, !!!dimension_labels)) %>%
  group_by(P3_7_label, dimension) %>%
  summarise(avg_val = mean(value, na.rm = TRUE), .groups = "drop") %>%
  drop_na()

# Plot heatmap with labels
ggplot(long_data, aes(x = dimension, y = fct_rev(P3_7_label), fill = avg_val)) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(limits = c(0, 5)) +
  labs(
    x = "Quality Dimension",
    y = "Healthcare Provider",
    fill = "Avg. Score",
    title = "Patient Experience Scores by Provider and Dimension"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text = element_text(size = 10),
    plot.title = element_text(face = "bold")
  )

```
```{r}
# ----------- RETURN PROPENSITY ---------------------
# Calculate return proportions
return_prop <- health_quality_data %>%
  filter(!is.na(P7_2), !is.na(P3_7_label)) %>%
  group_by(P3_7_label, P7_2) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(P3_7_label) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Extract % No (P7_2 == 0) per provider
no_order <- return_prop %>%
  filter(P7_2 == 0) %>%
  select(P3_7_label, no_prop = prop)

# Reorder factor levels based on descending % No
ordered_labels <- no_order %>%
  arrange(desc(no_prop)) %>%
  pull(P3_7_label)

# Apply factor levels to full data
return_prop <- return_prop %>%
  mutate(P3_7_label = factor(P3_7_label, levels = ordered_labels))

# Plot
ggplot(return_prop, aes(x = P3_7_label, y = prop, fill = factor(P7_2, levels = c(1, 0)))) +
  geom_col(position = "fill") +
  coord_flip() +
  scale_fill_manual(values = c("1" = "seagreen3", "0" = "firebrick3"), labels = c("1" = "Yes", "0" = "No")) +
  labs(
    x = "Healthcare Provider",
    y = "Proportion of Respondents",
    fill = "Would Return",
    title = "Willingness to Return by Provider (Ordered by % No Responses)"
  ) +
  theme_minimal()



# Reasons
# Create named labels for reasons
yes_labels <- c(
  P7_3_1 = "No other option",
  P7_3_2 = "Treated well",
  P7_3_3 = "Good care",
  P7_3_4 = "No payment",
  P7_3_5 = "Close to home",
  P7_3_6 = "Affordable",
  P7_3_7 = "Quick attention",
  P7_3_8 = "Received meds",
  P7_3_9 = "Short wait",
  P7_3_10 = "Clean facility",
  P7_3_11 = "Doctor explained",
  P7_3_12 = "Other",
  P7_3_13 = "Doesn’t know"
)

no_labels <- c(
  P7_4_1 = "Treated badly",
  P7_4_2 = "Disagreed with diagnosis",
  P7_4_3 = "Disagreed with treatment",
  P7_4_4 = "Did not improve",
  P7_4_5 = "Expensive",
  P7_4_6 = "Far from home",
  P7_4_7 = "No meds",
  P7_4_8 = "No materials",
  P7_4_9 = "Family not allowed",
  P7_4_10 = "Long wait",
  P7_4_11 = "Dirty facility",
  P7_4_12 = "Doctor didn’t explain",
  P7_4_13 = "Other",
  P7_4_14 = "Doesn’t know"
)

# Pivot long and tag reasons
reasons_long <- health_quality_data %>%
  select(starts_with("P7_3_"), starts_with("P7_4_")) %>%
  pivot_longer(
    everything(),
    names_to = "reason_code",
    values_to = "response"
  ) %>%
  filter(response == 1) %>%  # Only keep selected reasons
  mutate(
    type = case_when(
      str_detect(reason_code, "P7_3") ~ "Yes",
      str_detect(reason_code, "P7_4") ~ "No"
    ),
    reason = case_when(
      reason_code %in% names(yes_labels) ~ yes_labels[reason_code],
      reason_code %in% names(no_labels) ~ no_labels[reason_code],
      TRUE ~ "Unknown"
    )
  )

# Summarize counts
reason_summary <- reasons_long %>%
  group_by(type, reason) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(type) %>%
  mutate(prop = n / sum(n)) %>%
  ungroup()

# Plot
ggplot(reason_summary, aes(x = reorder(reason, prop), y = prop, fill = type)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~type, scales = "free_y") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(
    x = "Reason",
    y = "Proportion",
    title = "Reasons for Returning or Not Returning to Provider",
    subtitle = "Based on respondents' selected answers"
  ) +
  theme_minimal(base_size = 12)


## ---------- BEFORE STARTING: HEALTH CONDITIONS ------------

# Create condition mapping
condition_labels <- c(
  "1" = "Respiratory infections",
  "2" = "Pneumonia",
  "3" = "COPD",
  "4" = "Cough/Sore throat",
  "5" = "Tuberculosis",
  "6" = "Ear infection",
  "7" = "Conjunctivitis",
  "8" = "Asthma",
  "9" = "Allergies",
  "10" = "Heart disease",
  "11" = "Rheumatic fever",
  "12" = "Diabetes",
  "13" = "Hypertension",
  "14" = "Stroke",
  "15" = "Obesity",
  "16" = "Arthritis",
  "17" = "Diarrhea",
  "18" = "Gastritis/Ulcer",
  "19" = "Colitis",
  "20" = "Parasites",
  "21" = "Hepatitis",
  "22" = "Kidney disease",
  "23" = "UTI",
  "24" = "Exanthematous disease",
  "25" = "STI",
  "26" = "HIV/AIDS",
  "28" = "Dengue",
  "29" = "Poisoning",
  "30" = "Alcoholism",
  "32" = "Drug-related illness",
  "33" = "Accidental injury",
  "34" = "Injury from aggression",
  "36" = "Stress",
  "37" = "Depression",
  "38" = "Skin issues",
  "39" = "Oral disease",
  "40" = "Headache",
  "41" = "Fever (unspecified)",
  "42" = "Folk illnesses",
  "43" = "Pregnancy",
  "44" = "Cancer",
  "45" = "Other",
  "99" = "Don't know"
)

# Prepare the data - convert condition codes to factors with labels
health_quality_data <- health_quality_data %>%
  mutate(
    primary_condition_code = as.character(P1_2),
    primary_condition = factor(primary_condition_code, 
                               levels = names(condition_labels),
                               labels = condition_labels)
  )

# Rename provider
health_quality_data <- health_quality_data %>%
  mutate(provider = P3_7)

# Step 1: Chi-square test to see if conditions differ by provider
chi_result <- chisq.test(table(health_quality_data$primary_condition, health_quality_data$provider))
print(paste("Chi-square test result: X² =", round(chi_result$statistic, 2), 
            ", df =", chi_result$parameter, 
            ", p-value =", format.pval(chi_result$p.value, digits = 3)))

# Step 2: Calculate condition frequencies by provider
condition_by_provider <- health_quality_data %>%
  group_by(provider, primary_condition) %>%
  summarize(count = n(), .groups = "drop") %>%
  group_by(provider) %>%
  mutate(
    total = sum(count),
    percentage = count / total * 100
  ) %>%
  arrange(provider, desc(percentage))

# Step 3: Get the top 5 conditions for each provider
top_conditions_by_provider <- condition_by_provider %>%
  group_by(provider) %>%
  slice_max(order_by = percentage, n = 5) %>%
  ungroup()

# Step 4: Find conditions with the biggest variation across providers
condition_variation <- condition_by_provider %>%
  group_by(primary_condition) %>%
  summarize(
    max_pct = max(percentage, na.rm = TRUE),
    min_pct = min(percentage, na.rm = TRUE),
    variation = max_pct - min_pct,
    avg_pct = mean(percentage, na.rm = TRUE)
  ) %>%
  filter(!is.na(primary_condition)) %>%
  arrange(desc(variation))

top_varying_conditions <- condition_variation %>%
  slice_max(order_by = variation, n = 10) %>%
  pull(primary_condition)

# Step 5: Visualize the results

# Plot 1: Top 5 conditions by provider
plot1 <- ggplot(top_conditions_by_provider, 
                aes(x = reorder(primary_condition, percentage), y = percentage, fill = provider)) +
  geom_col() +
  facet_wrap(~ provider, scales = "free_y") +
  coord_flip() +
  theme_minimal() +
  labs(
    title = "Top 5 Primary Conditions by Provider Type",
    subtitle = "Percentage of patients seeking care for each condition",
    x = "Condition",
    y = "Percentage of Patients (%)"
  ) +
  theme(
    legend.position = "none",
    axis.text.y = element_text(size = 8)
  )

print(plot1)


```

```{r}
## ------------------ HYPOTHESIS 1: --------------------
## Without capitation or fee-for-service incentives and with low 
## monitoring, providers may induce less services. 



# Label providers
health_quality_data <- health_quality_data %>%
  mutate(
    provider = factor(provider, levels = c(1,2,3,4,5,6,7,8,9, 10, 11),  
                      labels = c("IMSS", "ISSSTE", "State-level ISSSTE",
                                 "Pemex", "Defensa", "Navy", "Seguro Popular", 
                                 "IMSS PROSPERA", "Other", "Private", "Other")),
    test_requested = ifelse(P6_1 == 1, 1, 0),
    sex = factor(SEXO, labels = c("Male", "Female")),
    medications = as.numeric(as.character(P5_1)),
    condition = as.factor(P1_2),  # self-reported condition
    age = as.numeric(EDAD)  
  )

health_quality_data$provider <- relevel(health_quality_data$provider, ref = "Private")

# Diagnostic tests requested
model_tests <- lm(test_requested ~ provider + age + sex + condition,
                   data = health_quality_data)

# Number of medications prescribed
model_meds <- lm(medications ~ provider + age + sex + condition,
                 data = health_quality_data)

# Summary with robust standard errors
coeftest(model_tests, vcov = vcovHC(model_tests, type = "HC1"))
coeftest(model_meds, vcov = vcovHC(model_meds, type = "HC1"))

##############################
#######  PLOTS FOR MODEL_TESTS
##############################
# -- FOR HEALTH CONDITIONS --
condition_map <- data.frame(
  term = paste0("condition", c(
    2, 3, 4, 5, 6, 7, 8, 9, 10, 11,
    12, 13, 14, 15, 16, 17, 18, 19, 20, 21,
    22, 23, 24, 25, 26, 28, 29, 30, 32, 33,
    34, 36, 37, 38, 39, 40, 41, 42, 43, 44,
    45, 99)
  ),
  condition = c(
    "Pneumonia", "COPD", "Cold/Sore throat", "Tuberculosis", "Ear infection", "Conjunctivitis", "Asthma", "Allergies", "Heart disease",
    "Rheumatic fever", "Diabetes", "Hypertension", "Stroke", "Obesity", "Arthritis", "Diarrhea", "Gastritis/Ulcer", "Colitis", "Parasites",
    "Hepatitis", "Kidney disease", "UTI", "Exanthematous disease", "STI", "HIV/AIDS", "Dengue", "Poisoning", "Alcoholism", "Drug-related illness",
    "Accidental injury", "Injury from aggression", "Stress", "Depression", "Skin issues", "Oral disease", "Headache", "Fever (unspecified)",
    "Folk illnesses", "Pregnancy", "Cancer", "Other", "Don't know"
  )
)

tidy_model <- tidy(model_tests, conf.int = TRUE)

# Filter only condition variables
tidy_conditions <- tidy_model %>%
  filter(grepl("^condition\\d+", term)) %>%
  left_join(condition_map, by = "term") %>%
  arrange(estimate)

# Create factor with ordering for plotting
tidy_conditions$condition <- factor(tidy_conditions$condition, levels = tidy_conditions$condition)


ggplot(tidy_conditions, aes(x = estimate, y = condition)) +
  geom_point(color = "steelblue", size = 3) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, color = "gray50") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Effect of Health Condition on Diagnostic Test",
    x = "Coefficient Estimate (with 95% CI)",
    y = "Condition"
  ) +
  theme_minimal(base_size = 13)


# -- NOW FOR PROVIDERS--
provider_map <- data.frame(
  term = c(
    "providerIMSS", "providerISSSTE", "providerState-level ISSSTE", 
    "providerPemex", "providerDefensa",
    "providerNavy", "providerSeguro Popular", "providerIMSS PROSPERA",
    "providerOther"
  ),
  provider = c(
    "Social Security (IMSS)", "ISSSTE", "State-level ISSSTE", "Pemex", "Defensa",
    "Navy", "Seguro Popular", "IMSS PROSPERA", "Other institution"
  )
)

# Tidy your model
tidy_model <- tidy(model_tests, conf.int = TRUE)

# Filter for provider terms only
tidy_providers <- tidy_model %>%
  filter(term %in% provider_map$term) %>%
  left_join(provider_map, by = "term") %>%
  arrange(estimate)

# Join the results on quality by provider
provider_summary <- provider_summary %>%
  rename(provider = P3_7_label)

tidy_providers <- tidy_providers %>%
  left_join(provider_summary, by = "provider")

# 1. Add significance stars
tidy_providers <- tidy_providers %>%
  mutate(
    stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*",
      p.value < 0.1 ~ ".",
      TRUE ~ ""
    ),
    coef_label = paste0(sprintf("%.2f", estimate), stars)
  )

# 2. Order by average quality (ensure provider is factor)
tidy_providers <- tidy_providers %>%
  mutate(provider = fct_reorder(provider, avg_quality))

# 3. Plot
ggplot(tidy_providers, aes(x = provider, y = estimate)) +
  geom_col(fill = "steelblue", width = 0.7) +
  geom_errorbar(aes(ymin = estimate - std.error,
                    ymax = estimate + std.error), width = 0.2) +
  geom_text(aes(label = coef_label), hjust = -0.1, size = 3.5) +
  coord_flip() +
  labs(
    title = "Effect of Healthcare Provider on Diagnostic Test Likelihood",
    x = "Provider",
    y = "Coefficient (with significance)"
  ) +
  theme_minimal()

plot_model(model_tests, show.values = TRUE, value.offset = 0.3) +
  coord_flip()


ggplot(tidy_providers, aes(x = estimate, y = provider)) +
  geom_point(color = "steelblue", size = 3) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, color = "gray50") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Effect of Provider Type on Number of Diagnostic Tests",
    subtitle = "Ordered by Mean Quality of Care",
    x = "Coefficient Estimate (with 95% CI)",
    y = "Provider (Highest to Lowest Mean Quality)"
  ) +
  theme_minimal(base_size = 13)

#############################
#######  PLOTS FOR MODEL_MEDS
#############################

# -- FOR HEALTH CONDITIONS --
condition_map <- data.frame(
  term = paste0("condition", c(
    2, 3, 4, 5, 6, 7, 8, 9, 10, 11,
    12, 13, 14, 15, 16, 17, 18, 19, 20, 21,
    22, 23, 24, 25, 26, 28, 29, 30, 32, 33,
    34, 36, 37, 38, 39, 40, 41, 42, 43, 44,
    45, 99)
  ),
  condition = c(
    "Pneumonia", "COPD", "Cold/Sore throat", "Tuberculosis", "Ear infection", "Conjunctivitis", "Asthma", "Allergies", "Heart disease",
    "Rheumatic fever", "Diabetes", "Hypertension", "Stroke", "Obesity", "Arthritis", "Diarrhea", "Gastritis/Ulcer", "Colitis", "Parasites",
    "Hepatitis", "Kidney disease", "UTI", "Exanthematous disease", "STI", "HIV/AIDS", "Dengue", "Poisoning", "Alcoholism", "Drug-related illness",
    "Accidental injury", "Injury from aggression", "Stress", "Depression", "Skin issues", "Oral disease", "Headache", "Fever (unspecified)",
    "Folk illnesses", "Pregnancy", "Cancer", "Other", "Don't know"
  )
)

tidy_model_meds <- tidy(model_meds, conf.int = TRUE)

# Filter only condition variables
tidy_conditions_meds <- tidy_model_meds %>%
  filter(grepl("^condition\\d+", term)) %>%
  left_join(condition_map, by = "term") %>%
  arrange(estimate)

# Create factor with ordering for plotting
tidy_conditions_meds$condition <- factor(tidy_conditions_meds$condition, levels = tidy_conditions_meds$condition)


ggplot(tidy_conditions_meds, aes(x = estimate, y = condition)) +
  geom_point(color = "steelblue", size = 3) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, color = "gray50") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Effect of Health Condition on Meds Prescription",
    x = "Coefficient Estimate (with 95% CI)",
    y = "Condition"
  ) +
  theme_minimal(base_size = 13)


# -- NOW FOR PROVIDERS--
provider_map <- data.frame(
  term = c(
    "providerIMSS", "providerISSSTE", "providerState-level ISSSTE", 
    "providerPemex", "providerDefensa",
    "providerNavy", "providerSeguro Popular", "providerIMSS PROSPERA",
    "providerOther"
  ),
  provider = c(
    "Social Security (IMSS)", "ISSSTE", "State-level ISSSTE", "Pemex", "Defensa",
    "Navy", "Seguro Popular", "IMSS PROSPERA", "Other institution"
  )
)

# Tidy your model
tidy_model_meds <- tidy(model_meds, conf.int = TRUE)

# Filter for provider terms only
tidy_providers_meds <- tidy_model_meds %>%
  filter(term %in% provider_map$term) %>%
  left_join(provider_map, by = "term") %>%
  arrange(estimate)

# Join the results on quality by provider
tidy_providers_meds <- tidy_providers_meds %>%
  left_join(provider_summary, by = "provider")

# 1. Add significance stars
tidy_providers_meds <- tidy_providers_meds %>%
  mutate(
    stars = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*",
      p.value < 0.1 ~ ".",
      TRUE ~ ""
    ),
    coef_label = paste0(sprintf("%.2f", estimate), stars)
  )

# 2. Order by average quality (ensure provider is factor)
tidy_providers_meds <- tidy_providers_meds %>%
  mutate(provider = fct_reorder(provider, avg_quality))

# 3. Plot
ggplot(tidy_providers_meds, aes(x = provider, y = estimate)) +
  geom_col(fill = "steelblue", width = 0.7) +
  geom_errorbar(aes(ymin = estimate - std.error,
                    ymax = estimate + std.error), width = 0.2) +
  geom_text(aes(label = coef_label), hjust = -0.1, size = 3.5) +
  coord_flip() +
  labs(
    title = "Effect of Healthcare Provider on # of Med Prescription",
    x = "Provider",
    y = "Coefficient (with significance)"
  ) +
  theme_minimal()

plot_model(model_tests, show.values = TRUE, value.offset = 0.3) +
  coord_flip()


ggplot(tidy_providers, aes(x = estimate, y = provider)) +
  geom_point(color = "steelblue", size = 3) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, color = "gray50") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Effect of Provider Type on Number of Med Prescription",
    subtitle = "Ordered by Mean Quality of Care",
    x = "Coefficient Estimate (with 95% CI)",
    y = "Provider (Highest to Lowest Mean Quality)"
  ) +
  theme_minimal(base_size = 13)

```

```{r}
## ------------------ HYPOTHESIS 2: --------------------
## Without capitation or fee-for-service incentives and with low 
## monitoring, providers put less effort per patient.

health_quality_data <- health_quality_data %>%
  # Consultation time in minutes
  mutate(consult_time = as.numeric(P4_7H) * 60 + as.numeric(P4_7M),
         
         # Effort perception variables (reverse coded so higher = more effort)
         ask_questions = ifelse(P8_5_2 %in% 1:4, 5 - as.numeric(P8_5_2), NA),
         enough_time = ifelse(P8_5_3 %in% 1:4, 5 - as.numeric(P8_5_3), NA),
         shared_decision = ifelse(P8_5_4 %in% 1:4, 5 - as.numeric(P8_5_4), NA),
         explain_clear = ifelse(P8_5_5 %in% 1:4, 5 - as.numeric(P8_5_5), NA),
         coordinate_specialist = ifelse(P8_5_6 %in% 1:4, 5 - as.numeric(P8_5_6), NA),
         
         # Other effort-relevant measures
         review_meds = ifelse(P8_8_1 == 1, 1, ifelse(P8_8_1 == 2, 0, NA)),
         explain_sidefx = ifelse(P8_8_2 == 1, 1, ifelse(P8_8_2 == 2, 0, NA)),
         
         # Provider recoding (if not yet done)
         provider = factor(P3_7, levels = c(1,2,3,4,5,6,7,8,9,10,11),
                           labels = c("IMSS", "ISSSTE", "State-level ISSSTE", "Pemex", "Defensa", "Navy", "Seguro Popular", 
                                      "IMSS PROSPERA", "Other", "Private", "Other")),
         
         sex = factor(SEXO, labels = c("Male", "Female")),
         age = as.numeric(EDAD)
  )

# Standardize variables
effort_vars <- c("consult_time", "ask_questions", "enough_time", "shared_decision",
                 "explain_clear", "coordinate_specialist", "review_meds", "explain_sidefx")

health_quality_data <- health_quality_data %>%
  mutate(across(all_of(effort_vars), ~ scale(.)[,1], .names = "z_{col}")) %>%
  rowwise() %>%
  mutate(effort_index = mean(c_across(starts_with("z_")), na.rm = TRUE)) %>%
  ungroup()

# Define the base comparison for the Private
health_quality_data$provider <- relevel(health_quality_data$provider, ref = "Private")

# Define the model
effort_model <- lm(effort_index ~ provider + age + sex, data = health_quality_data)
effort_model_conditions <- lm(effort_index ~ provider + age + sex + condition, data = health_quality_data)


# Print the summary
summary(effort_model)
summary(effort_model_conditions)


## -- FOR PROVIDERS --

pred_df <- health_quality_data %>%
  group_by(provider) %>%
  summarise(mean_effort = mean(effort_index, na.rm = TRUE),
            se = sd(effort_index, na.rm = TRUE)/sqrt(n()))

# First get the original plot with sorting
plot <- plot_model(effort_model, 
                   type = "est", 
                   sort.est = TRUE,  # Sort by effect size
                   title = "Effects on Provider Effort Index",
                   show.values = TRUE,
                   show.p = TRUE,
                   value.offset = 0.3,
                   colors = c("firebrick", "steelblue"))  # Red for negative, blue for positive

# Extract and modify the data while preserving order
plot_data <- plot$data
# Store original order
plot_data$original_order <- 1:nrow(plot_data)
# Remove "provider" from labels
plot_data$term <- gsub("provider", "", plot_data$term)

# Rebuild plot with modified labels but preserved order
ggplot(plot_data, aes(x = reorder(term, original_order), y = estimate, color = estimate > 0)) +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high), fatten = 1) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  scale_color_manual(values = c("firebrick", "steelblue")) +
  geom_text(aes(label = paste0(sprintf("%.2f", estimate), 
                               ifelse(p.value < 0.001, "***", 
                                      ifelse(p.value < 0.01, "**",
                                             ifelse(p.value < 0.05, "*", ""))))),
            vjust = -0.5, size = 3.5) +
  labs(title = "Effects on Provider Effort Index",
       y = "Estimate",
       x = "") +
  theme_sjplot() +
  theme(axis.text.y = element_text(size = 10),
        plot.title = element_text(face = "bold", size = 14),
        legend.position = "none") +
  coord_flip()


# -- FOR HEALTH CONDITIONS --
condition_map <- data.frame(
  term = paste0("condition", c(
    2, 3, 4, 5, 6, 7, 8, 9, 10, 11,
    12, 13, 14, 15, 16, 17, 18, 19, 20, 21,
    22, 23, 24, 25, 26, 28, 29, 30, 32, 33,
    34, 36, 37, 38, 39, 40, 41, 42, 43, 44,
    45, 99)
  ),
  condition = c(
    "Pneumonia", "COPD", "Cold/Sore throat", "Tuberculosis", "Ear infection", "Conjunctivitis", "Asthma", "Allergies", "Heart disease",
    "Rheumatic fever", "Diabetes", "Hypertension", "Stroke", "Obesity", "Arthritis", "Diarrhea", "Gastritis/Ulcer", "Colitis", "Parasites",
    "Hepatitis", "Kidney disease", "UTI", "Exanthematous disease", "STI", "HIV/AIDS", "Dengue", "Poisoning", "Alcoholism", "Drug-related illness",
    "Accidental injury", "Injury from aggression", "Stress", "Depression", "Skin issues", "Oral disease", "Headache", "Fever (unspecified)",
    "Folk illnesses", "Pregnancy", "Cancer", "Other", "Don't know"
  )
)

tidy_model_effort <- tidy(effort_model_conditions, conf.int = TRUE)

# Filter only condition variables
tidy_model_effort <- tidy_model_effort %>%
  filter(grepl("^condition\\d+", term)) %>%
  left_join(condition_map, by = "term") %>%
  arrange(estimate)

# Create factor with ordering for plotting
tidy_model_effort$condition <- factor(tidy_model_effort$condition, levels = tidy_model_effort$condition)


ggplot(tidy_model_effort, aes(x = estimate, y = condition)) +
  geom_point(color = "steelblue", size = 3) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, color = "gray50") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Effect of Health Condition on Effort",
    x = "Coefficient Estimate (with 95% CI)",
    y = "Condition"
  ) +
  theme_minimal(base_size = 13)

```
## Hypothesis 3
```{r}
## ----------------- HYPOTHESIS 3 --------------------
## Moral hazard leads to inefficient care, not better care.

# Outcome: would return (Yes = 1)
health_quality_data <- health_quality_data %>%
  mutate(
    health_outcome = ifelse(P7_1 %in% c(1,2), 1, 0)
  )

health_quality_data <- health_quality_data %>%
  mutate(
    num_meds = P5_1
  )

health_quality_data <- health_quality_data %>%
  filter(!is.na(provider))    # remove any missing‐provider rows

# 1) Ensure provider is coded and labeled
health_quality_data <- health_quality_data %>%
  mutate(
    provider = factor(provider, levels = unique(provider)),  # preserve existing order
    system   = recode(as.character(provider),
                      "1"  = "Social Security (IMSS)",
                      "2"  = "ISSSTE",
                      "3"  = "State-level ISSSTE",
                      "4"  = "Pemex",
                      "5"  = "Defensa",
                      "6"  = "Navy",
                      "7"  = "Seguro Popular",
                      "8"  = "IMSS PROSPERA",
                      "9"  = "Private insurance",
                      "10" = "Other institution",
                      "11" = "Not affiliated",
                      "99" = "Doesn't know")
  )

# 2) TEST_REQUESTED: run reg within each provider, pull slope + CI
test_df <- health_quality_data %>%
  group_by(provider, system) %>%
  do(tidy(
    lm(health_outcome ~ test_requested + age + sex + factor(primary_condition), data = .),
    conf.int = TRUE
  )) %>%
  filter(term == "test_requested") %>%
  mutate(original_order = row_number())

# 3) NUM_MEDS: same but for num_meds
meds_df <- health_quality_data %>%
  group_by(provider, system) %>%
  do(tidy(
    lm(health_outcome ~ num_meds + age + sex + factor(primary_condition), data = .),
    conf.int = TRUE
  )) %>%
  filter(term == "num_meds") %>%
  mutate(original_order = row_number())

# 4) Function to plot any of these
plot_by_system <- function(df, title_text) {
  ggplot(df,
         aes(x = reorder(system, original_order),
             y = estimate,
             color = estimate > 0)) +
    geom_pointrange(aes(ymin = conf.low, ymax = conf.high),
                    fatten = 1) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
    scale_color_manual(values = c("firebrick", "steelblue")) +
    geom_text(aes(label = paste0(
                     sprintf("%.2f", estimate),
                     ifelse(p.value < 0.001, "***",
                       ifelse(p.value < 0.01, "**",
                              ifelse(p.value < 0.05, "*", "")
                       )
                     )
                   )),
              vjust = -0.5, size = 3.5) +
    labs(title = title_text,
         y     = "Slope (95% CI)",
         x     = "") +
    theme_sjplot() +
    theme(axis.text.y     = element_text(size = 10),
          plot.title      = element_text(face = "bold", size = 14),
          legend.position = "none") +
    coord_flip()
}

# 5) Draw them
plot_by_system(test_df, "Effect of Test Requested on Better Care, by Provider")
plot_by_system(meds_df, "Effect of Number of Medications on Better Care, by Provider")

```

```{r}
# Total medical spending
health_quality_data <- health_quality_data %>%
  mutate(
    lab_cost = as.numeric(P6_4),
    med_cost = as.numeric(P5_6),
    other_cost = as.numeric(P6_6_1) + as.numeric(P6_6_2) + as.numeric(P6_6_3),
    total_cost = rowSums(across(c(lab_cost, med_cost, other_cost)), na.rm = TRUE)
  )

# Regression: Costs ~ Treatment Intensity

# 2) Extract test_requested slopes by provider
test_df <- health_quality_data %>%
  group_by(provider, system) %>%
  do(tidy(
    lm(total_cost ~ test_requested + age + sex + factor(primary_condition), data = .),
    conf.int = TRUE
  )) %>%
  filter(term == "test_requested") %>%
  mutate(original_order = row_number())

# 3) Extract num_meds slopes by provider
meds_df <- health_quality_data %>%
  group_by(provider, system) %>%
  do(tidy(
    lm(total_cost ~ num_meds + age + sex + factor(primary_condition), data = .),
    conf.int = TRUE
  )) %>%
  filter(term == "num_meds") %>%
  mutate(original_order = row_number())

# 4) Plotting function
plot_by_system <- function(df, title_text) {
  ggplot(df,
         aes(x = reorder(system, original_order),
             y = estimate,
             color = estimate > 0)) +
    geom_pointrange(aes(ymin = conf.low, ymax = conf.high),
                    fatten = 1) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
    scale_color_manual(values = c("firebrick", "steelblue")) +
    geom_text(aes(label = paste0(
                     sprintf("%.2f", estimate),
                     ifelse(p.value < 0.001, "***",
                       ifelse(p.value < 0.01, "**",
                              ifelse(p.value < 0.05, "*", "")
                       )
                     )
                   )),
              vjust = -0.5, size = 3.5) +
    labs(title = title_text,
         y     = "Slope (95% CI)",
         x     = "") +
    theme_sjplot() +
    theme(axis.text.y     = element_text(size = 10),
          plot.title      = element_text(face = "bold", size = 14),
          legend.position = "none") +
    coord_flip()
}

# 5) Draw the two coefficient plots
plot_by_system(test_df, "Effect of Test Requested on Total Cost, by System")
plot_by_system(meds_df, "Effect of Number of Medications on Total Cost, by System")
```

## Hypothesis 4
Wait time and access
```{r}
health_quality_data <- health_quality_data %>%
  # convert factors/characters to numeric
  mutate(
    wait_time_rating    = as.numeric(as.character(P4_6)),
    consult_minutes     = as.numeric(P4_7H) * 60 + as.numeric(P4_7M),
    access_apt_rating   = as.numeric(as.character(P8_2)),
    access_after_hours  = as.numeric(as.character(P8_3)),
    access_by_phone     = as.numeric(as.character(P8_4))
  )

# Quick check:
health_quality_data %>%
  select(wait_time_rating, consult_minutes,
         access_apt_rating, access_after_hours, access_by_phone) %>%
  summary()


```

```{r}
# Wait time and access
# 1) Compute wait_time_score & consult_minutes
health_quality_data <- health_quality_data %>%
  mutate(
    wait_time_rating  = as.numeric(as.character(P4_6)),
    wait_time_score   = 6 - wait_time_rating,               # higher = worse wait
    consult_minutes   = as.numeric(P4_7H) * 60 + as.numeric(P4_7M),
    provider          = factor(P3_7,
                               levels = c(9,1,2,3,4,5,6,7,8,10,11),  # put 9="Private" first
                               labels = c(
                                 "Private insurance",
                                 "Social Security (IMSS)","ISSSTE","State-level ISSSTE",
                                 "Pemex","Defensa","Navy","Seguro Popular",
                                 "IMSS PROSPERA","Other institution","Not affiliated"
                               )),
    primary_condition = factor(P1_2),
    sex               = factor(SEXO, labels = c("Male","Female")),
    age               = as.numeric(EDAD)
  )

# 2) Model 1: Wait time
mod_wait <- lm(wait_time_score ~ provider + age + sex + factor(primary_condition),
               data = health_quality_data)

wait_coefs <- tidy(mod_wait, conf.int = TRUE) %>%
  filter(grepl("^provider", term)) %>%
  # drop the reference level (intercept) row if it sneaks in
  filter(term != "(Intercept)") %>%
  mutate(
    system         = sub("^provider", "", term),
    system         = gsub("\\.", " ", system),  # clean up label if needed
    original_order = row_number()
  )

# 3) Model 2: Consultation minutes
mod_consult <- lm(consult_minutes ~ provider + age + sex + factor(primary_condition),
                  data = health_quality_data)

consult_coefs <- tidy(mod_consult, conf.int = TRUE) %>%
  filter(grepl("^provider", term)) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    system         = sub("^provider", "", term),
    system         = gsub("\\.", " ", system),
    original_order = row_number()
  )

# 4) Plotting function
plot_coef <- function(df, title_text) {
  ggplot(df, aes(x = reorder(system, original_order),
                 y = estimate,
                 color = estimate > 0)) +
    geom_pointrange(aes(ymin = conf.low, ymax = conf.high),
                    fatten = 1) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
    scale_color_manual(values = c("firebrick", "steelblue")) +
    geom_text(aes(label = paste0(
                     sprintf("%.2f", estimate),
                     ifelse(p.value < 0.001, "***",
                       ifelse(p.value < 0.01, "**",
                         ifelse(p.value < 0.05, "*", "")
                       )
                     )
                   )),
              vjust = -0.5, size = 3.5) +
    labs(title = title_text,
         y     = "Difference vs. Private  (95% CI)",
         x     = "") +
    theme_sjplot() +
    theme(axis.text.y     = element_text(size = 10),
          plot.title      = element_text(face = "bold", size = 14),
          legend.position = "none") +
    coord_flip()
}

# 5) Draw the two plots
plot_coef(wait_coefs,    "Difference in Wait Time (versus Private)")
plot_coef(consult_coefs, "Difference in Consult Minutes (versus Private)")


```






